Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-04-05 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
New version of figure with incomplete dataset

```{r patterns, fig.height=3, fig.width=4.1, fig.cap = "Missingness in multilevel data", echo=FALSE}
plot_md_set()
```

@@

```{r imp_impact}
imp <- mice(impact_NA, method = meth, predictorMatrix = pred, m = 2, maxit = 1)
# imp_impact <- impact_NA %>%
#   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
#   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
# look at jomo for categorical variables?
# semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
# pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
# make assumptions of these methods explicit!
```
```{r impact-fit}
fit <- imp_impact_pmm %>% 
  with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
tidy(pool(fit))
as.mitml.result(fit)
# testEstimates(as.mitml.result(fit))
```

>>>>

```{r patterns, fig.height=2.5, fig.width=4.1, fig.cap = "Missingness in multilevel data", echo=FALSE}
# blankcol <- c(rep("", 4), "...", "")
# dat <- expand.grid(rows = 1:6, cols = 1:6) %>% 
#   cbind(text = c("1", "1", "2", "2", "...", "N", rep(blankcol, 3), rep("...", 6), blankcol),
#         miss = c(rep("", 14), "NA", "NA", "", "", "NA", "", "", "NA", rep("", 14)))
dat <- expand.grid(rows = 1:6, cols = 1:6) %>% 
  cbind(text = c("1", "1", "2", "2", "", "N", rep("", 30)),
        miss = c(rep("", 14), "NA", "NA", "", "", "NA", "", "", "NA", rep("", 14)))

ggplot(dat, aes(x = cols, y = rows)) +
  geom_tile(fill = "white", color = "black") +
  geom_text(aes(label = text), color = "black") +
  geom_text(aes(label = miss), color = mice:::mdc(2), family = "mono", fontface = "bold") + 
  scale_x_continuous(breaks = 1:6, labels = c("cluster", expression(X[1]), expression(X[2]), expression(X[3]), "...", expression(X[p])), name = NULL, position = "top") +
  scale_y_continuous(breaks = 1:6, labels = c(1:4, "...", "n"), name = NULL, trans = "reverse") +
  theme_minimal() +
  theme(panel.grid = element_blank())

@@

```{r imp_impact, eval = FALSE}
imp <- mice(impact_NA, method = meth, predictorMatrix = pred, m = 2, maxit = 1)
# imp_impact <- impact_NA %>%
#   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
#   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
# look at jomo for categorical variables?
# semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
# pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
# make assumptions of these methods explicit!
```
```{r impact-fit, eval = FALSE}
fit <- imp_impact_pmm %>% 
  with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
tidy(pool(fit))
as.mitml.result(fit)
# testEstimates(as.mitml.result(fit))
```

### 2022-04-05 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
New version of figure with incomplete dataset

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/impact_md-1} \end{center}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/impact_md-2} \end{center}

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 22 x 5
   term                 estimate std.error statistic  p.value
   <chr>                   <dbl>     <dbl>     <dbl>    <dbl>
 1 (Intercept)           -3.37     0.205     -16.4   1.97e-60
 2 typeRCT                0.395    0.177       2.23  2.59e- 2
 3 age                    0.0301   0.00191    15.8   5.01e-56
 4 motor_score3          -0.618    0.0829     -7.45  9.33e-14
 5 motor_score4          -1.00     0.0850    -11.8   6.07e-32
 6 motor_score5/6        -1.53     0.0874    -17.5   7.92e-69
 7 pupil                  0.434    0.0398     10.9   9.36e-28
 8 ct                     0.428    0.0344     12.4   1.50e-35
 9 as.factor(name)CSTAT  -0.177    0.199      -0.892 3.72e- 1
10 as.factor(name)EBIC    0.560    0.174       3.22  1.30e- 3
# ... with 12 more rows
\end{CodeOutput}
\end{CodeChunk}

If we run the multilevel model on the incomplete date, we get a warning
about unindentifyability. This means that with CCA as missing data
method, we cannot trust the estimates. To still obtain some estimates to
compare later, we fit a simplified analysis model: using the cluster
variable as indicator in the model.

\begin{CodeChunk}
\begin{CodeInput}
R> pred <- quickpred(impact_NA)
R> pred[pred == 1] <- 2
R> pred[, "name"] <- -2
R> pred["mort", ] <- 2
R> pred[, "mort"] <- 2
R> diag(pred) <- 0
R> pred[c("name", "type", "age", "motor_score", "mort"), ] <- 0
R> pred
\end{CodeInput}
\begin{CodeOutput}
            name type age motor_score pupil ct hypox hypots tsah edh mort
name           0    0   0           0     0  0     0      0    0   0    0
type           0    0   0           0     0  0     0      0    0   0    0
age            0    0   0           0     0  0     0      0    0   0    0
motor_score    0    0   0           0     0  0     0      0    0   0    0
pupil         -2    0   0           2     0  2     0      2    0   0    2
ct            -2    2   2           2     2  0     0      0    2   2    2
hypox         -2    2   0           2     0  0     0      2    0   0    2
hypots        -2    0   0           2     2  0     2      0    0   0    2
tsah          -2    0   2           0     0  2     0      0    0   0    2
edh           -2    0   0           0     0  2     0      0    0   0    2
mort           0    0   0           0     0  0     0      0    0   0    0
\end{CodeOutput}
\begin{CodeInput}
R> meth <- make.method(impact_NA)
R> meth[meth != ""] <- "2l.pmm"
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> # imp_impact <- impact_NA %>%
R> #   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
R> #   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
R> # look at jomo for categorical variables?
R> # semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
R> # pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
R> # make assumptions of these methods explicit!
\end{CodeInput}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 9 x 7
  effect   group term                    estimate std.error statistic   p.value
  <chr>    <chr> <chr>                      <dbl>     <dbl>     <dbl>     <dbl>
1 fixed    <NA>  (Intercept)              -2.66     0.194      -13.7   1.11e-42
2 fixed    <NA>  typeRCT                  -0.412    0.172       -2.40  1.64e- 2
3 fixed    <NA>  age                       0.0298   0.00190     15.7   2.56e-55
4 fixed    <NA>  as.factor(motor_score)2  -0.631    0.0823      -7.66  1.82e-14
5 fixed    <NA>  as.factor(motor_score)3  -0.994    0.0845     -11.8   5.61e-32
6 fixed    <NA>  as.factor(motor_score)4  -1.54     0.0868     -17.7   2.88e-70
7 fixed    <NA>  pupil                     0.438    0.0397      11.0   3.01e-28
8 fixed    <NA>  ct                        0.430    0.0343      12.5   5.34e-36
9 ran_pars name  sd__(Intercept)           0.257   NA           NA    NA       
\end{CodeOutput}
\end{CodeChunk}

We should impute the variables \texttt{ct} and \texttt{pupil} and any
auxiliary variables we might want to use to impute these incomplete
analysis model variables. We can evaluate which variables may be useful
auxiliaries by plotting the pairwise complete correlations:

\begin{CodeChunk}
\begin{CodeInput}
R> plot_corr(impact_NA, label = T)
\end{CodeInput}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/impact_corr-1} \end{center}

@@

This shows us that \texttt{hypox} and \texttt{hypot} would not be useful
auxiliary variables for imputing \texttt{ct}. Depending on the minimum
required correlation, \texttt{tsah} could be useful, while \texttt{edh}
has the strongest correlation with \texttt{ct} out of all the variables
in the data and should definitely be included in the imputation model.
For the imputation of \texttt{pupil}, none of the potential auxiliary
variables has a very strong relation, but \texttt{hypots} could be used.
We conclude that we can exclude \texttt{hypox} from the data since this
is neither an analysis model variable, nor an auxiliary variable for
imputation.

\begin{CodeChunk}
\begin{CodeInput}
R> impact_NA <- select(impact_NA, -c(hypox, hypots))
R> pred <- quickpred(impact_NA)
R> plot_pred(pred)
\end{CodeInput}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/impact-pred-1} \end{center}

\begin{CodeInput}
R> pred[pred == 1] <- 2
R> pred["mort", ] <- 2
R> pred[, "mort"] <- 2
R> pred[c("name", "type", "age", "motor_score", "mort"), ] <- 0
R> pred[, "name"] <- -2
R> diag(pred) <- 0
R> plot_pred(pred)
\end{CodeInput}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/impact-pred-2} \end{center}

\begin{CodeInput}
R> meth <- make.method(impact_NA)
R> meth[meth != ""] <- "2l.pmm"
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> imp <- mice(impact_NA, method = meth, predictorMatrix = pred, m = 2, maxit = 1)
R> # imp_impact <- impact_NA %>%
R> #   mutate(name = as.integer(name), motor_score = as.numeric(motor_score)) %>%
R> #   mice::mice(., m = 2, maxit = 1, method = meth, predictorMatrix = pred)
R> # look at jomo for categorical variables?
R> # semi-cont with jomo is not ideal (schafer, '97) because you need 2-step approach
R> # pmm is better (more efficient) because it will still look for donors (maybe outside of cluster) based on predictive distance, even for very small clusters
R> # make assumptions of these methods explicit!
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- imp_impact_pmm %>% 
+   with(glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial")) 
R> tidy(pool(fit))
R> as.mitml.result(fit)
R> # testEstimates(as.mitml.result(fit))
\end{CodeInput}
\end{CodeChunk}

Compare estimates to complete data

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mort ~ type + age + as.factor(motor_score) + pupil + ct + (1 | name), family = "binomial", data = impact) 
R> tidy(fit)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 7
   effect   group term                   estimate std.error statistic    p.value
   <chr>    <chr> <chr>                     <dbl>     <dbl>     <dbl>      <dbl>
 1 fixed    <NA>  (Intercept)             -2.29     0.171      -13.4   4.94e- 41
 2 fixed    <NA>  typeRCT                 -0.338    0.172       -1.96  5.01e-  2
 3 fixed    <NA>  age                      0.0322   0.00160     20.2   2.13e- 90
 4 fixed    <NA>  as.factor(motor_score~  -0.593    0.0698      -8.50  1.95e- 17
 5 fixed    <NA>  as.factor(motor_score~  -0.904    0.0713     -12.7   8.43e- 37
 6 fixed    <NA>  as.factor(motor_score~  -1.29     0.0732     -17.6   1.67e- 69
 7 fixed    <NA>  pupilNone                1.32     0.0616      21.4   2.09e-101
 8 fixed    <NA>  pupilOne                 0.626    0.0711       8.80  1.36e- 18
 9 fixed    <NA>  ctIII                    0.810    0.0700      11.6   5.12e- 31
10 fixed    <NA>  ctIV/V                   0.832    0.0593      14.0   9.47e- 45
11 ran_pars name  sd__(Intercept)          0.277   NA           NA    NA        
\end{CodeOutput}
\end{CodeChunk}
